---
title: Bowsher Rx 
output:
  tufte::tufte_handout: default
params:
  drug: Tylenol
  max: 60
  min: 20
resource_files:
- template/template.pptx
- .Renviron
- .httr-oauth
- template/template2.pptx
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(flexdashboard)
library(stringr)
library(bigrquery)
library(openfda)
library(dplyr)
library(dbplyr)
library(DBI)
library(ggplot2)
library(blastula)
library(ggthemes)
library(formatR)

con <- dbConnect(
  bigrquery::bigquery(),
  project = 'bigquery-public-data',
  dataset = 'nlm_rxnorm',
  billing = bq_test_project()
)
pathways <- tbl(con, 'rxn_all_pathways_01_18')

# Get all brand name drugs from BigQuery
all_names <- pathways %>% 
  filter(TARGET_TTY == 'BN') %>% 
  select(TARGET_NAME) %>% 
  group_by(TARGET_NAME) %>% 
  summarise() %>% 
  collect()

get_adverse <- function(gender, brand_name, age) {
  fda_query("/drug/event.json") %>%
    fda_filter("patient.drug.openfda.brand_name", brand_name) %>% 
    fda_filter("patient.patientsex", gender) %>% 
    fda_filter("patient.patientonsetage", age) %>% 
    fda_count("patient.reaction.reactionmeddrapt.exact") %>% 
    fda_limit(10) %>% 
    fda_exec()
}


create_age <- function(min, max){#
  sprintf('[%d+TO+%d]', min, max)
}
```

# Adverse Event Data for `r params$drug` 

```{r datacheck}
if (!(params$drug %in% all_names$TARGET_NAME)) {
  print(sprintf("WARNING: Drug %s not found in FDA database.", params$drug))
  knitr::opts_chunk$set(eval = FALSE)
}
```

Report run as of `r Sys.Date()`. This report is limited to data on events for individuals aged `r params$min` to `r params$max`. `r params$drug` has the following active ingredients:

```{r ingredients}
ingredients <-  pathways %>%
  filter(TARGET_NAME == params$drug, TARGET_TTY == 'BN', SOURCE_TTY %in% c('IN', 'SY')) %>% 
  select(SOURCE_NAME, SOURCE_TTY) %>% 
  collect()

knitr::kable(ingredients %>% select(`Active Ingredients` = SOURCE_NAME))
```


`r tufte::margin_note("Active Ingredients from Big Query's rxnorm dataset.")`

This report pulls data from the openFDA API. The following outputs summarize the data pulls and indicate that the event data was fetched correctly.

```{r eventdata, echo=FALSE, message=FALSE, warning=FALSE}
age <- create_age(params$min, params$max)

jnk <- capture.output(male <- get_adverse("1", params$drug, age))
if (!is.null(male)) {
  male$gender = 'male'
}

jnk <-  capture.output(female <-get_adverse("2", params$drug, age))
if (!is.null(female)) {
  female$gender = 'female'
}

adverse <- rbind(male, female)

if(is.null(adverse)) {
  print(sprintf("WARNING: No adverse event data found for %s  in openFDA API.", params$drug))
  knitr::opts_chunk$set(eval = FALSE)
}
```

With the data in hand, we can analyze the events themselves. This report includes a basic overview.

## All Events

```{r fig.fullwidth = TRUE}
events <- adverse %>% 
  group_by(term) %>% 
  summarise(count = sum(count)) %>% 
  ggplot() +
    geom_bar(aes(reorder(term,count), count), stat = 'identity') +
    coord_flip() +
    labs(
      title = params$drug,
      x = NULL,
      y = NULL
    ) +
    theme_minimal()
events
```

## Events by Gender

In addition to looking at all the events, it is informative to see if the events are distributed differently by gender. ^[A thorough statistical analysis would be necessary to conclude that the genders differ significantly.] The following chart provides a useful exploratory analysis of the differences between gender. Note that some events were not experience by both genders, and some events were experienced but not to the same severity. Keep in mind the age range that this plot reflects.

```{r}
gender <- ggplot(adverse) +
    geom_bar(aes(reorder(term,count), count, fill = gender), stat = 'identity') +
    facet_wrap(~gender)+
    coord_flip() +
    labs(
      title = params$drug,
      x = NULL,
      y = NULL
    ) +
    theme_minimal() + 
    guides(fill = FALSE) + 
    scale_fill_manual(values = c("#d54a30","#4c83b6"))
gender
```
^[Age Range is `r params$min`  - `r params$max`. Effects may vary with different ages. ]


Finally, this report can also be used to generate a summary PPT and email. Press the email icon in the upper right hand corner to send yourself the results.


```{r ppt, message=FALSE}
callr::r_copycat(
  function(events, gender, event_data, drug){
    rmarkdown::render('template/template.Rmd', 
                      output_file = 'events.pptx',
                      quiet = TRUE)
  },
  list(events, gender, adverse, params$drug)
)

file.copy('template/events.pptx', './events.pptx')
subject <- sprintf('%s Adverse Event Report ', params$drug)

email <- blastula::compose_email(
  body = "
  Hi Team,
  
  Attached is your PPT summary of the adverse events for {params$drug}. The main adverse events are highlighted below. Data as of {Sys.Date()}.
  
  
  {add_ggplot(events, width = 6, height = 6)}
  
  
  Thanks,
  
  Phil
  
  
  Copyright 2018 Bowsher Rx
  "
)
```

```{r email, message=FALSE}
rmarkdown::output_metadata$set(
  rsc_email_subject = subject,
  rsc_email_body_html = email$html_str,
  rsc_email_images = email$images,
  rsc_email_attachments = list("events.pptx")
)
```




